---
title: Comparison of time efficiency of a same function in R
author: ~
date: '2019-05-28'
slug: comparison-of-time-efficiency-of-a-same-function-in-r
categories: ["R","Julia","CUDA"]
tags: ["R","CUDA", "GPU","Julia","Byte code compilation"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

Hi guys,

in this Post it is verified the efficiency in question of time of execution of functions created for Matrix multiplication.

An array with dimensions $1000\times1000$ was generated using the 
`rnorm` function in R.

1. The first function called `mult.matrix` uses the base function of R for matrix multiplication, for example, consider A as a square matrix, multiplication is obtained by `A%*%A`.
2. In the second function, the byte code compiler which can
increase the speed of some code was used. Using the compiler is an easy way to get improvements in speed.
3. The third function uses the graphical processing unit (GPU) as the processing unit.
NVIDIA GPU uses CUDA programming language, however it is
integration with software R is possible. Some functions are implemented in the `gpuR` package, one of them is matrix creation using the GPU, in this way it is possible to realize the multiplication.
4.The fourth function analyzed is using the R and Julia integration.
Through the `JuliaCall` package it is possible to integrate R with Julia, and thus send and return functions and results.

```{r}
# Required Packages
library(tidyverse)
library(microbenchmark)
library(gpuR) # GPU Package
library(JuliaCall) # Julia Package for Integration

# Opening the integration of R with Julia 
julia <- julia_setup()

# Function for matrix multiplication with base function of R
mult.matrix <- function( x ){
  matrix(x, nrow=1000) %*%
    matrix(x, nrow=1000)
}

# Function for matrix multiplication with Byte code Compilation in base function of R 
mult.matrix.byte <- compiler::cmpfun(mult.matrix)


# Function for matrix multiplication with GPU
gpu.mult.matrix <- function( x ){
  vclMatrix(matrix(x, nrow=1000), type="float") %*%
    vclMatrix(matrix(x, nrow=1000), type="float")
}

# Function for matrix multiplication with Julia
mult.matrix.julia <- function(x){
  
  # Send Matrix
  julia_assign("x", matrix(x, nrow=1000)); 
  # multiplication
  julia_command("y = x*x;")
  # return for R
  res <- julia_eval("y")
  return(res)
}

#Create database
x <- rnorm(1000*1000)

#Using the microbenchmark function
result <- microbenchmark(
  mult.matrix(x),
  mult.matrix.byte(x),
  gpu.mult.matrix(x),
  mult.matrix.julia(x)
)

print(result)

autoplot(result)

```

The results show that the most efficient method are obtained with the function that uses the GPU to calculate the multiplication.
The basic function of R, such as the byte code compilation, presented results similar, this show that compilation not performance gain in the speed of this specific function.
Lastly, the less efficient method was the function that  integrate R and Julia. This may have happened, for the time needed to send the data to Julia and in the end get results return.

Although the example is simple, such as matrix multiplication, these calculation alternatives can be applied to other functions and thus gain performance gains in your applications.


for more details see:

[NVIDIA CUDA](https://developer.nvidia.com/nvidia-developer-zone)

[Li, Changcheng. "JuliaCall: an R package for seamless integration between R and Julia."](https://www.theoj.org/joss-papers/joss.01284/10.21105.joss.01284.pdf)

[Wickham, Hadley. Advanced r. Chapman and Hall/CRC, 2014.](http://adv-r.had.co.nz/)

