---
title: Web Scrapping introduction with R
author: ~
date: '2019-12-18'
slug: web-scrapping-introduction-with-r
categories: ["R","Web Scrapping"]
tags: ["R","Web Scrapping"]
---

Hi,
in this post I'll talk about Web Scraping with R.

The Project "SERENATA DE AMOR" [https://serenata.ai/](https://serenata.ai/) is an open project that uses data science - the same technologies used by giants like Google, Facebook and Netflix - for the purpose of monitoring public spending and sharing information in a way accessible to everyone.

The project has already reported 629 suspicious reimbursement to the Chamber of Deputies involving 216 different congresspeople and more than 378k BRL.

This amount corresponds to 371% of what was raised through our crowdfunding campaign (~US$ 25,000.00)

Do you want  to know how the report system works? [You can read more in this 
post (in 
Portuguese)](https://medium.com/data-science-brigade/como-est%C3%A1-acontecendo-a-hackaton-de-den%C3%BAncias-da-opera%C3%A7%C3%A3o-serenata-de-amor-a8bd193e0c76).

The purpose of this Post is to collect information on websites (in this case in the 
[Jarbas](https://jarbas.serenata.ai/dashboard/chamber_of_deputies/reimbursement/) )
and do analyzes with the data in R.
Here was just made a chart,
with some deputies and the reimbursement amounts requested by them.


Step 1: Reading the URL.

To read the html page, we will use the "read_html" of the package *rvest*.

```{r}
library(rvest)
url <- "https://jarbas.serenata.ai/dashboard/chamber_of_deputies/reimbursement/"
jarbas_webpage <- read_html(url)
```

Step 2: Now, the best part of the rvest library is that you can extract the data from html nodes, what that means is you can straightaway select the nodes with their id’s or css classes and extract the text from the html tags. So i went to my url and fired up the "*firebug*" on the browser and soon figuered that the names of the hikes have been encapsulated in the "*.field-congressperson_name*" css class, using this css class i can extract all the trail names on the webpage.

There are 2 functions that we will use here:

1. html_nodes : Use this function to extract the nodes that we like (in this case nodes with ".field-congressperson_name" as css class
2. html_text: Use this function to extract the text in between the html nodes (in this case our deputies names)

```{r}
#Scraping Trail Names using css class ‘field-congressperson_name’
jarbas_names_html <-html_nodes(jarbas_webpage, '.field-congressperson_name')
jarbas_names <- html_text(jarbas_names_html)
head(jarbas_names)
```

Similarly, now I will do this for all other attributes: Subquota Translated, Provider. Each of these attributes have their own css classes: field-subquota_translated, field-supplier_info.


```{r}
#SUBQUOTA TRANSLATED
jarbas_subquota_html <-html_nodes(jarbas_webpage, '.field-subquota_translated')
jarbas_subquota <- html_text(jarbas_subquota_html)
head(jarbas_subquota)

#Provider
jarbas_provider_html <-html_nodes(jarbas_webpage, '.field-supplier_info')
jarbas_provider <- html_text(jarbas_provider_html)
head(jarbas_provider)
```


One thing to note, when we extract value of reimbursement, it is in the form of an character: Eg: R\$ 139,76, R\$ 40,23. We want it in numeric form, To convert it into a numeric form, used a library: "*Stringr*" and the function:`str_extract`

the script below was used to convert it to numeric form:

```{r}
library(stringr)
#value in brazilian real
jarbas_value_html <-html_nodes(jarbas_webpage, '.field-value')
jarbas_value <- html_text(jarbas_value_html)
head(jarbas_value)
#R$ 139,76" "R$ 40,23"  "R$ 72,05"  "R$ 72,05"  "R$ 56,39"  "R$ 235,64"  
#Extracted data is in the form of character, we need to extract digits and convert it into numeric format
jarbas_value <- as.numeric(sub(",",".",str_extract(jarbas_value,pattern = "\\-*\\d+,\\s{0,}\\d+")))
head(jarbas_value,25)
```

Step 3: Now that we have all our information lets collate it into a dataframe.
To make the result visible, only the first name of each deputies has been selected.

```{r}
#Combining all the extracted features of the trails
jarbas_names <- str_extract(jarbas_names,pattern = boundary("word"))
jarbas_df <- data.frame(
  Name = jarbas_names,
  Subquota = jarbas_subquota,
  Provider = jarbas_provider,
  Value = jarbas_value
  )
str(jarbas_df)
```
  
Step 4:  As an early analysis, we can use graph between the reimbursement amount and the name of deputies.
The first 50 rows of the data were selected so that the chart in the Post was visible.
For this part, using a library ggplot2 to visualise the data

```{r, fig.width=10}
library(ggplot2)

jarbas_df <- jarbas_df[1:50,]

ggplot(
  jarbas_df, aes(Value,Name,colour=Subquota)) +
  geom_point() +
  labs(title="", x ="reimbursement",
       y = "deputies",colour="SUBQUOTA TRANSLATED")
```



