---
title: k-NN classification in R and Julia
author: Jodavid Ferreira
date: '2019-05-29'
slug: k-nn-classification-in-r-and-julia
categories: ["Classification","R","Julia"]
tags: ["R","Julia","classification"]
---

Hi guys,

in this post will be presented an efficient algorithm and one of the 
most basic in the area of statistical classification, this algorithm is the $k$-NN ($k$ Nearest Neighbors).


The k-Nearest Neighbors algorithm was proposed by Cover
e Hart and is considered as a non-parametric classifier.
Given a test sample where the class is
unknown, the method finds the nearest $k$ in the training
set for each observation of the test sample over
a distance and assigns a class to the observation
according to the majority vote of the classes of these
neighbors.

The algorithm $k$-NN estimates the
probability of an observation
belonging to each group based
on the nearest neighbor $k$
information defined as follows.
Consider $G$ non-empty
partitions, which are
$C_1,C_2, \ldots,C_G$, such that,
$C_j \subset \mathbb{R}^p$
is a subset
of size $n_j$, for
$j = 1, \ldots,g$,
$n = \sum_{j=1}^{n} n_j$,
$C_i\cap C_j = \emptyset$ for all
$i \neq j.$
Let $\bm{y} \in \mathbb{R}^p$
an observation to be classified and
$\{\bm{x}_{i_1},\bm{x}_{i_2},\ldots,\bm{x}_{i_k} \}$ the set of $k$ neighbors, such that,
$k\leq n$ and
$\{{i_1},{i_2},\ldots,{i_k} \} \subset \{1,2,\ldots,n\}$.
So for $j = 1, \ldots,G$,
the classification rule of the $k$-NN
is defined as:

$$
\widehat{c}_{_{kNN}} = \mathop{\text{ arg max}}_{c} \widehat{P}_k(G=c),
$$

\noindent where $\widehat{P}_k(G=c) = k^{-1}\sum_{\nu=1}^{k}\mathbb{I}_c(\bm{x}_{i_\nu})$ and

$$
\mathbb{I}_c(\mathbf{x} _{i_\nu}) = \left\lbrace \begin{array}{rl}
1,&\text{if } \mathbf{x}_{i_\nu} \in C_j,\\
0,& \text{if } \mathbf{x}_{i_\nu} \notin C_j.
\end{array}
\right.
$$
To illustrate our example we will use the 'Iris' database. Below is a graph of two variables of the 'Iris' database, this base contains 4 variables that we will use in the classification and the fifth variable is the 'true class' of the observations.

```{r}
library(plotly)

p <- ggplot(iris,aes(Sepal.Length,Petal.Length,colour=Species)) +
  geom_point()

ggplotly(p)

```


Let's divide our experiment into 3 steps:

1. Take a random sample from the Iris database to represent our 
training sample;
2. The sample not selected for the training will be considered the
test sample, with this sample we will classifier using the algorithm 
$k$-NN;
3. As we know the true class of the observations of the test sample, it is possible to obtain the classification accuracy and verify the efficiency of the algorithm $k$ -NN.

Let's repeat the same steps for the R and Julia languages.

#### In R:

```{r}
# required packages
library(class)
#database
head(iris)

# training sample
set.seed(1) # seed it is necessary for obtained same results
sample.iris <- sample(1:nrow(iris),50) # 50 observations
sample.iris
training <- iris[sample.iris,-ncol(iris)]
class_training <- iris[sample.iris,ncol(iris)]
test <- iris[-sample.iris,-ncol(iris)]
class_test <- iris[-sample.iris,ncol(iris)]

# k-NN with 3 Neighbour
resul <- knn(training,test,class_training,k=3)

# classification accuracy - R
mean(resul==class_test)
```



#### In Julia:

```{julia}
using StatsKit
using RDatasets

#<> ----------------------------------------- <>#
# k-NN Functions
## majority vote
function maj_vote(yn)
    ## majority vote
    cm = countmap(yn)
    mv = -999
    lab = nothing
    tot = 1e-8
    
    for (k,v) in cm
        tot += v
        
          if v > mv
            mv = v
            lab = k
          end
    
    end
    
    prop = /(mv, tot)
    return [lab, prop]
end

## KNN label prediction
function knn(x, y, x_new , k)
    x = convert(Array, x)
    y = convert(Array, y)
    x_new = convert(Array, x_new)
    n,p = size(x)
    n2,p2 = size(x_new)
    ynew = DataFrame(Class=String[], 
        maj=Float64[])
    for i in 1:n2 ## over new x_new
        res = zeros(n,2)
        for j in 1:n ## over x
            ## euclidean distance between rows - each row is a subject
            res[j,:] = [j , euclidean(x[j,:], x_new[i,:])]
        end
        ## sort rows by distance and index - smallest distances
        res2 = sortslices(res, dims = 1, by = x -> (x[2], x[1]))
        ## get indices for the largest k distances
        ind = convert(Array{Int}, res2[1:k, 1])
        ## take majority vote for the associated indices
        res3 = maj_vote(y[ind])
        ynew = vcat(ynew,DataFrame(Class=res3[1], maj = res3[2] ))
    end
    ## return the predicted labels
    return ynew
end

#<> ----------------------------------------- <>#

iris = dataset("datasets", "iris")
sample_iris = [40,56,85,134,30,131,137,95,90,9,29,25,143,53,105,68,97,132,51,102,122,28,84,16,34,49,2,48,107,42,58,72,59,22,96,77,91,13,82,46,114,71,148,60,57,83,3,50,75,70]
training = iris[sample_iris,1:size(iris,2).!=ncol(iris)]
class_training = iris[sample_iris,ncol(iris)]
sample_test_iris = [1,4,5,6,7,8,10,11,12,14,15,17,18,19,20,21,23,24,26,27,31,32,33,35,36,37,38,39,41,43,44,45,47,52,54,55,61,62,63,64,65,66,67,69,73,74,76,78,79,80,81,86,87,88,89,92,93,94,98,99,100,101,103,104,106,108,109,110,111,112,113,115,116,117,118,119,120,121,123,124,125,126,127,128,129,130,133,135,136,138,139,140,141,142,144,145,146,147,149,150]
test = iris[sample_test_iris,1:size(iris,2).!=ncol(iris)]
class_test = iris[sample_test_iris,ncol(iris)]

# k-NN with 3 Neighbour
predict_k3 = knn(training,class_training,test,3)

# classification accuracy - Julia
class_accuracy = mean(predict_k3[:,1] .== class_test)
```

The experiment was performed using a training sample with 50 observations and a test sample with 100 observations.
In the R software, the $k$-NN algorithm is in the `class` library, in Julia it was necessary to write the $k$-NN code, since it was not found  nothing library with the method.
As was to be expected, using the same observations for the
training sample and test sample, the R and the Julia obtained the same results, that was 97% accuracy.